---
title: "1.wrangling-yield"
output: html_document
---


# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)

#install.packages("knitr")
#install.packages("viridis")

library(viridis)
```

```{r reading data}
yield16 <- read_csv("../data/Deer Run1 2016 Yield.csv")

yield16
```
# Exploratory data analysis
```{r class}
class(yield16)
```

```{r summary}
summary(yield16) 
```


```{r yield distribution - density plot}
ggplot(yield16, aes(x=DryYield))+
  geom_density()+
  geom_rug()
```

```{r speed distribution - density plot}
ggplot(yield16, aes(x=speed_mph))+
  geom_density()+
  geom_rug()

```

```{r yield map}
ggplot(yield16, aes(x=Long,
                    y=Lat))+
  geom_point(aes(color=DryYield))+
  scale_color_viridis()
  scale_color_gradient(low="red", high = "green")

```

```{r yield restricted map}
ggplot(filter(yield16, DryYield<5000) , aes(x=Long,
                                             y=Lat))+
  geom_point(aes(color=DryYield))+
  scale_color_viridis()
  scale_color_gradient(low="red", high = "green")

```

```{r yield x speed}
ggplot(yield16, aes(x=speed_mph, 
                    y=DryYield))+
  geom_point()+
  geom_smooth()

```


```{r data needs}
yield16

```


# Wrangling  
```{r yield16_w1 - clean_names}
# Standardizing names
yield16_w1 <- clean_names(yield16)


yield16_w1
```

```{r yield16_w2 - select}
# Selecting needed columns
yield16_w2 <-   dplyr::select(yield16_w1,
                              speed_mph,
                              dry_yield,
                              lat,
                             long)


yield16_w2
```


```{r yield16_w3 - rename}
# Rename to include units
yield16_w3 <- rename(yield16_w2,
                     yield_lbac=dry_yield,
                     speed_mph=speed_mph)

yield16_w3

```

```{r yield16_w4 - mutate}
# Creating yield and speed in SI
yield16_w4 <- mutate(yield16_w3,
                     speed_kmh=speed_mph*1.60934,
                     yield_kgha=yield_lbac*0.453592/0.404686
)

yield16_w4


```

```{r yield16_w5 - st_as_sf}
# Creating spatial object
yield16_w5 <- st_as_sf(yield16_w4, 
                       coords = c("long", "lat"))

yield16_w5

class(yield16_w5)
```



```{r yield16_w6 - st_set_crs}
# Assigning native coordinate reference system EPSG identifier
yield16_w6 <- st_set_crs(yield16_w5,
                         value = 4326) #WGS84

yield16_w6
```



```{r yield16_w7 - st_transform}
# Transforming from unprojected to projected (UTM NAD83 16 n)
yield16_w7 <- st_transform(yield16_w6,
                           crs = 6345)

yield16_w7
```
 

```{r yield16_w - piping it all}
yield16_w <- yield16 %>% 
  clean_names() %>%
  dplyr::select(speed_mph,
                dry_yield,
                lat,
               long) %>%
  rename(speed_mph=speed_mph,
         yield_lbac=dry_yield) %>%
  mutate(yield_kgha=yield_lbac*0.453592/0.404686,
         speed_kmh=speed_mph*1.60934, yield_buac=yield_lbac*0.0167) %>%
  st_as_sf(coords = c("long", "lat")) %>%
  st_set_crs(value = 4326) %>%
  st_transform(crs = 6345)

yield16_w


```
```{r raw SD}

print(sd(yield16_w$yield_buac))

```


# Yield Editor  
```{r yield map raw}
yieldmap_raw <- ggplot(yield16_w)+
  geom_sf(aes(color=yield_lbac), size=.1)+
  scale_color_viridis()
  scale_color_gradient(low="red", high = "green")

yieldmap_raw
```



## Min/max yield
```{r min max yield - density}
ggplot(data=yield16_w, aes(x=yield_lbac))+
  geom_density()+
  geom_rug()

```

```{r yield summary}
summary(yield16_w$yield_buac)
```


```{r yield quantiles}
lowq_yield <- quantile(yield16_w$yield_lbac, 0.1)
lowq_yield

upq_yield <- quantile(yield16_w$yield_lbac, 0.99)
upq_yield

iqr <- upq_yield - lowq_yield
iqr

miny <- lowq_yield - (0.25*iqr)
miny

maxy <- upq_yield + (0.25*iqr)
maxy
```

```{r min max yield annotated- density}
ggplot()+
  geom_density(data=yield16_w, aes(x=yield_lbac))+
  geom_vline(xintercept = lowq_yield, color="red", linetype=2)+
  geom_vline(xintercept = upq_yield, color="blue", linetype=2)+
  geom_vline(xintercept = miny, color="red", linetype=1)+
  geom_vline(xintercept = maxy, color="blue", linetype=1)+
  annotate(geom="rect",
           ymin=0, ymax=Inf,
           xmin=miny, xmax=maxy,
           alpha=.6, 
           fill="green")

```


```{r yield filtering}
yield16_w_yf <- yield16_w %>%
  filter(yield_lbac > miny & yield_lbac < maxy)

yield16_w_yf
```
Before filtering we had a total of 73,574 data rows.  
How many rows now after filtering?  73,190


```{r yield filtered - stats}
summary(yield16_w_yf$yield_buac) 

print(sd(yield16_w_yf$yield_buac))
```


## Min/max speed  
```{r min max speed - density}
ggplot(data=yield16_w, aes(x=speed_mph))+
  geom_density()+
  geom_rug()

```

```{r speed summary}
summary(yield16_w$speed_mph)
```


```{r speed quantiles}
lowq_speed <- quantile(yield16_w$speed_mph, 0.08)
lowq_speed

upq_speed <- quantile(yield16_w$speed_mph, 0.99)
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (0.18*iqr)
minv

maxv <- upq_speed + (0.18*iqr)
maxv
```


```{r min max speed annotated- density}
ggplot()+
  geom_density(data=yield16_w, aes(x=speed_mph))+
  geom_vline(xintercept = lowq_speed, color="red", linetype=2)+
  geom_vline(xintercept = upq_speed, color="blue", linetype=2)+
  geom_vline(xintercept = minv, color="red", linetype=1)+
  geom_vline(xintercept = maxv, color="blue", linetype=1)+
  annotate(geom="rect",
           ymin=0, ymax=Inf,
           xmin=minv, xmax=maxv,
           alpha=.6, 
           fill="green")


```

```{r speed filtering}
yield16_w_sf <- yield16_w_yf %>%
  filter(speed_mph > minv & speed_mph < maxv)

yield16_w_sf
```
Before filtering we had a total of 73,190 data rows.  
How many rows now after filtering?  68,358


```{r speed filtered - stats}
summary(yield16_w_sf$speed_mph) 
```

# Position  

```{r importing boundary}

boundary <- read_sf("..//data/boundary/DR1_boundary.geojson")  

boundary

```

```{r boundary_w}
boundary_w <- boundary %>%
  st_transform(crs = 6345)
  
boundary_w
```

```{r boundary_w plot}
yieldmap_raw+ 
  geom_sf(data=boundary_w, fill=NA)

```

Let's create a negative buffer of 35-m.  
```{r buffer}
buffer <- boundary_w %>%
  st_buffer(dist = -35)

buffer
```

```{r buffer plot}
yieldmap_raw+ 
  geom_sf(data=boundary_w, fill=NA)+
  geom_sf(data=buffer, color="orange", fill=NA)

```

```{r position filtering}
yield16_w_pf <- yield16_w_sf %>%
  st_intersection(buffer)

yield16_w_pf
```
Before filtering we had a total of 68,358 data rows.  
How many rows now after filtering? 58,778 

```{r position summary}
summary(yield16_w_pf$yield_lbac)
```

# Raw vs Cleaned data  
## Summary statistics
```{r raw summary}
yield16_w %>%
  summarise(n=length(yield_buac),
            min=min(yield_buac),
            mean=mean(yield_buac),
            max=max(yield_buac),
            sd=sd(yield_buac),
            cv=(sd/mean)*100
            )
```

```{r filtered summary}
yield16_w_pf %>%
  summarise(n=length(yield_buac),
            min=min(yield_buac),
            mean=mean(yield_buac),
            max=max(yield_buac),
            sd=sd(yield_buac),
            cv=(sd/mean)*100
            )
```

## Density plot  
```{r raw vs clean density}
ggplot()+
  geom_density(data=yield16_w, 
               aes(x=yield_lbac), 
               fill="blue", alpha=.5)+
  geom_density(data=yield16_w_pf, 
               aes(x=yield_lbac), 
               fill="yellow", alpha=.5)

```

# Map  
```{r raw map}
yieldmap_raw
```

```{r cleaned map}
yieldmap_clean <- ggplot(yield16_w_pf)+
  geom_sf(aes(color=yield_lbac), size=.1)+
  scale_color_viridis()

yieldmap_clean
```

```{r side-by-side maps}
#install.packages("patchwork")
library(patchwork)

yieldmap_raw + yieldmap_clean

# saving to file
ggsave("../output/rawvsclean16.png", width = 10, height = 6)
```

# Exporting cleaned version  
```{r exporting clean}
write_sf(yield16_w_pf,"../output/Rojas_Yield16_Clean.geojson")
```





